---
title: "index"
author: "Jan Helfenstein"
format: html
editor: visual
date: 2025-03-24
---

## Index

## packages

```{r}
library(tidyverse)
library(readr)
library(ggplot2)
library(gt)
library(knitr)
```

## Get data

```{r}
survey <- read_rds(here::here("data/processed/survey-data-processed.rds"))
```

## Visualize data

The following point plot @fig-rating displays when survey was taken and how the users rated ETH's recycling situation. It is possible that the given rating reflected the tidyness at ETH at that time.

```{r}
#| label: fig-rating
#| fig-cap: "Survey timeline and ETH's recycling rating"
survey_vis <- survey |> 
  rename("recycling_frequency" = recycler_cat) |> 
  group_by(weekday) |> 
  mutate(amount = n())

ggplot(data = survey_vis, 
       mapping = aes(x = datetime,
                     y = rating
                    )
       ) + 
  geom_point(shape = 2, size = 3) + 
  labs(x = "survey timestamp")

```

The next bar plot @fig-recycling shows the relation between the participants' recycling frequency and their rating of ETH's recycling situation. Generally, people who recycle more often tend to give a higher rating.

```{r}
#| label: fig-recycling
#| fig-cap: "Rating of ETH's recycling situation by groups with different recycling frequencies"
ggplot(data = survey_vis, 
       mapping = aes(x = rating,
                     fill = recycling_frequency
                    )
       ) + 
  geom_bar(width = 0.2) +
  labs(title = "Rating of ETH's recycling situation",
       subtitle = "answers from 7 students",
       fill = "recycling frequency")

?labs
```

The following pie chart visualizes the recycling frequency:

```{r}
# summarize the data
# count the occurences of frequency category, calculate percentages
survey_pie_freq <- survey_vis |> 
  group_by(recycling_frequency) |> 
  summarize(n = n(),
            fraction = n()/nrow(survey_vis))

# create pie chart
# had help by perplexity.ai, link:
# https://www.perplexity.ai/search/how-to-create-a-pie-chart-with-f0XcFVCEQYOZ7Mmn3crgeA
pie(survey_pie_freq$fraction, labels = survey_pie_freq$recycling_frequency, main = "How often do ETH students recycle their waste? n = 7")


```

The following two pie chart visualizes how many people have a walkable station nearby, and how many people walk to a recycling station to recycle their waste.

```{r}
survey_pie_walkable <- survey_vis |> 
  group_by(walkable_station) |> 
  summarize(n = n(),
            fraction = n()/nrow(survey_vis))

pie(survey_pie_walkable$fraction, labels = survey_pie_walkable$walkable_station, main = "Do ETH students have a recycling station in walkable distance? n = 7")
```

```{r}
# create dataframe only with people with station in walkable distancee
survey_walkable <- survey_vis |> 
  filter(walkable_station == "TRUE")

# create summary of this df for pie chart
survey_pie_walker <- survey_walkable |> 
  group_by(regular_walker) |> 
  summarize(n = n(),
            fraction = n()/nrow(survey_walkable)
            )

# create pie chart
pie(survey_pie_walker$fraction, labels = survey_pie_walker$regular_walker, main = "How many of these ETH students walk to the nearby station? n = 5")
```

It is satisfying to see that all 5 students who have a walkable station nearby will walk there to recycle their waste.

How many of these selected people are regular recyclers?

```{r}
survey_pie_freq2 <- survey_walkable |> 
  filter(regular_walker == "TRUE") |> 
  group_by(recycling_frequency) |> 
  summarize(n = n(),
            fraction = n()/nrow(survey_walkable)
            )
  
# create pie chart
pie(survey_pie_freq2$fraction, labels = survey_pie_freq2$recycling_frequency, main = "How often do these regular walkers recycle their goods? n = 5")
```

## Table of the data with gt()

The following table sumarizes the relation between recycling frequency and the rating that students give to ETH's recycling situation:

```{r}

survey_table_recycler <- survey |> 
  group_by(recycler_cat) |> 
  summarise(count = n(),
            mean = mean(rating),
            sd = sd(rating),
            min = min(rating),
            max = max(rating)
           )

# create quick table with gt
survey_table_recycler |> 
  gt() |> 
  tab_header(title = "Rating of recycling at ETH by students, grouped by recycling frequency",
             subtitle = "Data from 7 students in 2025") |> 
  cols_label(recycler_cat = "recycling_frequency",
             mean = "mean_rating")

```

## Table of the data with kable()

In @tbl-survey-recycler , the same numbers are displayed, showing the recycling frequency and how it corresponds to ETH's recycling rating. But this time, the table was generated with the kable() function.

```{r}
#| label: tbl-survey-recycler
#| tbl-cap: "Rating of recycling at ETH by students, grouped by recycling frequency. Data from 6 students."
survey_table_recycler |> 
  rename(`recycling_frequency` = recycler_cat,
         `mean_rating` = mean) |> 
  kable()
```

@tbl-survey-recycler shows that people who always recycle their waste are rating the recycling situation at ETH higher than others. People who recycle less often tend to rate the recycling situation lower, at two or three stars.

This is also visualized in @fig-recycling\

## Saving the dataframes used for visualization

saving it in data/final folder\

```{r}
# dataframe for point plot
write_csv(survey_vis, here::here("data/final/survey_vis.csv"))

# dataframe for pie chart
write_csv(survey_pie_freq, here::here("data/final/survey_pie_freq.csv"))
write_csv(survey_pie_walkable, here::here("data/final/survey_pie_walkable.csv"))
write_csv(survey_pie_walker, here::here("data/final/survey_pie_walker.csv"))
write_csv(survey_pie_freq2, here::here("data/final/survey_pie_freq2.csv"))

# dataframe for table
write_csv(survey_table_recycler, here::here("data/final/survey_table_recycler.csv"))

```

\
