---
title: "index"
author: "Jan Helfenstein"
format: html
editor: visual
date: 2025-03-24
---

## Index

## packages

```{r}
library(tidyverse)
library(readr)
library(ggplot2)
library(gt)
library(knitr)
```

## Get data

```{r}
survey <- read_rds(here::here("data/processed/survey-data-processed.rds"))
```

## Visualize data

The following point plot @fig-rating displays when survey was taken and how the users rated ETH's recycling situation. It is possible that the given rating reflected the tidyness at ETH at that time.

```{r}
#| label: fig-rating
#| fig-cap: "Survey timeline and ETH's recycling rating"
survey_vis <- survey |> 
  rename("recycling_frequency" = recycler_cat) |> 
  group_by(weekday) |> 
  mutate(amount = n())

ggplot(data = survey_vis, 
       mapping = aes(x = datetime,
                     y = rating
                    )
       ) + 
  geom_point(shape = 2, size = 3)

```

The next bar plot @fig-recycling shows the relation between the participants' recycling frequency and their rating of ETH's recycling situation. Generally, people who recycle more often tend to give a higher rating.

```{r}
#| label: fig-recycling
#| fig-cap: "Rating of ETH's recycling situation by groups with different recycling frequencies"
ggplot(data = survey_vis, 
       mapping = aes(x = rating,
                     fill = recycling_frequency
                    )
       ) + 
  geom_bar(width = 0.2)

```

The following pie chart visualizes the recycling frequency:

```{r}
# summarize the data
# count the occurences of frequency category, calculate percentages
survey_pie_freq <- survey_vis |> 
  group_by(recycling_frequency) |> 
  summarize(n = n(),
            fraction = n()/nrow(survey_vis))

# create pie chart
# had help by perplexity.ai, link:
# https://www.perplexity.ai/search/how-to-create-a-pie-chart-with-f0XcFVCEQYOZ7Mmn3crgeA
pie(survey_pie_freq$fraction, labels = survey_pie_freq$recycling_frequency, main = "How often do ETH students recycle their waste? n = 6")


# what did not work:
#  mutate(f_count = count(case_when(recycling_frequency == "always", ...))
  
# f_count <- as.data.frame(table(survey_vis$recycling_frequency))
  
# mutate(table_freq = table(recycling_frequency))

# mutate(percent_freq = 100*recycling_frequency / sum(recycling_frequency))

#pie(survey_vis$recycling_frequency, main = "Pie Chart on recycling frequeny")
```

The following two pie chart visualizes how many people have a walkable station nearby, and how many people walk to a recycling station to recycle their waste.

```{r}
survey_pie_walkable <- survey_vis |> 
  group_by(walkable_station) |> 
  summarize(n = n(),
            fraction = n()/nrow(survey_vis))

pie(survey_pie_walkable$fraction, labels = survey_pie_walkable$walkable_station, main = "Do ETH students have a recycling station in walkable distance? n = 6")
```

```{r}
# create dataframe only with people with station in walkable distancee
survey_walkable <- survey_vis |> 
  filter(walkable_station == "TRUE")

# create summary of this df for pie chart
survey_pie_walker <- survey_walkable |> 
  group_by(regular_walker) |> 
  summarize(n = n(),
            fraction = n()/nrow(survey_walkable)
            )

# create pie chart
pie(survey_pie_walker$fraction, labels = survey_pie_walker$regular_walker, main = "How many of these ETH students walk to the nearby station? n = 4")
```

It is satisfying to see that all students who have a walkable station nearby will walk there to recycle their waste.

How many of these selected people are regular recyclers?

```{r}
survey_pie_freq2 <- survey_walkable |> 
  filter(regular_walker == "TRUE") |> 
  group_by(recycling_frequency) |> 
  summarize(n = n(),
            fraction = n()/nrow(survey_walkable)
            )
  
# create pie chart
pie(survey_pie_freq2$fraction, labels = survey_pie_freq2$recycling_frequency, main = "How often do these regular walkers recycle their goods? n = 4")
```

## Table of the data with gt()

The following table sumarizes the relation between recycling frequency and the rating that students give to ETH's recycling situation:

```{r}

survey_table_recycler <- survey |> 
  group_by(recycler_cat) |> 
  summarise(count = n(),
            mean = mean(rating),
            sd = sd(rating),
            min = min(rating),
            max = max(rating)
           )

# create quick table with gt
survey_table_recycler |> 
  gt() |> 
  tab_header(title = "Rating of recycling at ETH by students, grouped by recycling frequency",
             subtitle = "Data from 6 students in 2025") |> 
  cols_label(recycler_cat = "recycling_frequency",
             mean = "mean_rating")

```

## Table of the data with kable()

In @tbl-survey-recycler , the same numbers are displayed, showing the recycling frequency and how it corresponds to ETH's recycling rating. But this time, the table was generated with the kable() function.

```{r}
#| label: tbl-survey-recycler
#| tbl-cap: "Rating of recycling at ETH by students, grouped by recycling frequency. Data from 6 students."
survey_table_recycler |> 
  rename(`recycling_frequency` = recycler_cat,
         `mean_rating` = mean) |> 
  kable()
```

@tbl-survey-recycler shows that people who always recycle their waste are rating the recycling situation at ETH higher than others. People who recycle "most of the times" rate the recycling situation 1 point lower, at 3 stars.

This is also visualized in @fig-recycling\
\
\
